Universal relations after GW170817

### 2018-11-28 paper/bin-love-after-gw170817.tex
resolving conflict

\definecolor{red(ncs)}{rgb}{0.77, 0.01, 0.2}
\newcommand{\ny}[1]{\textcolor{blue}{NY: #1} }
\newcommand{\kc}[1]{\textcolor{red}{KC: #1} }

@@

Fig.~\ref{fig:binLove} shows the improved binary Love universal relation for 3 different mass ratios: $q=0.9$, $q=0.75$, and $q=0.5$ for both constrained and unconstrained sets of EoSs.
We observe that the maximum recorded EoS variation under the constrained set of EoSs is $23.65\%$, $9.76\%$, and $2.15\%$ for mass ratios of $q=0.9$, $q=0.75$, and $q=0.50$, respectively.
This shows considerable improvement from the universalities of $\sim 50\%$, $\sim 20\%$, and $\sim 3\%$ found in Ref.~\cite{Yagi:binLove}.
In addition, the unconstrained set of EoSs show similar universality to that recorded in Ref.~\cite{Yagi:binLove}.
\begin{figure*}
\begin{center} 
\includegraphics[width=\textwidth]{binLove.eps}
\end{center}
\caption{
Binary Love universal relations shown for the constrained EoSs (left) and unconstrained EoSs (right).
In this figure, the dashed black line corresponds to the fit given by Eq.~\ref{eq:binLovefit}, while the dotted blue, dashed magenta, and solid green lines represent the 35 EoSs for mass ratios of $q=0.9$, $q=0.75$, and $q=0.50$, respectively.
The maximal EoS variation for each value of $q$ is $23.65\%$, $9.76\%$, and $2.15\%$ respectively, compared to $\sim 50\%$, $\sim 20\%$, and $\sim 3\%$ found in Ref.~\cite{Yagi:binLove}.
These values are comparable to the maximal EoS variation for unconstrained EoSs as expected, found to be $41.66\%$, $12.64\%$, and $3.83\%$.
Additionally shown is the binary Love relations for hybrid stars (dash-dash-dotted curves) as discussed in Sec.~\label{sec:hybrid}.
Observe the large deviations from the fit as the mass ratio $q$ increases, returning near the fit before the transitional pressure at large values of $\Lambda_s$.
}
\label{fig:binLove}
\end{figure*}

@@

\subsubsection{Hybrid quark-hadron stars}\label{sec:hybrid}
\begin{figure}
\begin{center} 
\includegraphics[width=\linewidth]{hybridML.eps}
\end{center}
\caption{
$M-\tilde{\Lambda}$ relations for the ACB5 model hybrid star. 
Note the strong phase transition into a quark-matter branch at $\sim 1.4 \text{ M}_{\odot}$.
As described in the text, HS/HS and NS/NS binaries have individual masses chosen such that both stars are contained within one branch -- resulting in high $q$ values.
Constrastingly, HS/NS binaries have individual masses chosen from each branch resulting in small $q-values$.
}
\label{fig:hybridML}
\end{figure} 
Transitional quark-hadron matter stars undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at a given mass transition.
We focus on the ACB5 model shown in Fig.~\ref{fig:hybridML}, which observes a transition mass at $\sim 1.4 \text{ M}_{\odot}$~\cite{Paschalidis2018}.
This results in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq 1.4 \text{ M}_{\odot}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote this as ``HS"), and (ii) small-mass ($M \leq 1.4 \text{ M}_{\odot}$) hadronic stars with no internal transition to quark matter (we denote this as ``NS").
We consider 3 classes of binaries:
\begin{itemize}
\item HS/NS binary: consists of one massive hybrid star, and one small hadronic neutron star,
\item NS/NS binary: consists of two small-mass hadronic neutron stars,
\item HS/HS binary: consists of two massive hybrid stars.
\end{itemize}

@@

To determine how each of class of binary NSs obey our improved binary Love universal relations, we pick a representative set of 10 binary pairs from each class.
Binary Love values are then computed for each pair, then residuals from the binary Love curve found in Sec.~\ref{sec:nuclear} are determined and compared to that from the constrained set of EoSs.
The results are shown in Fig.~\ref{fig:hybrid}.
First, we observe that the NS/NS combination of stars gives consistent results to that found earlier -- as expected.
Next, we see that HS/HS combinations agree moderately with previous relations\footnote{Note that these EoSs do not fit within GW170817's 90\% posterior on EoS - so we excpect deviations consistent with the ``unconstrained" set of EoSs, rather than the ``constraind" set.}, while HS/NS combinations disagree significantly.
This can be explained by the large reduction in tidal deformability for hybrid stars, are shown in Fig. 2 of Ref.~\cite{Paschalidis2018}: when one star lies on the hadronic branch while the other is on the quark branch, they each see vast differences in $\tilde{\Lambda}$ - disrupting the binary Love relations as found for stars on the hadronic branch.
In addition, the discrepancy can be seen in Fig.~\ref{fig:binLove} where the binary Love relations for the ACB5 EoS is shown in relation to the constrained nuclear matter EoSs.
Observe how the residuals from the fit are quite good for low pressures (large $\Lambda_s$) when the the binaries consist of a NS/NS combination, before reaching the transitional pressure when the deviations become increasingly large for HS/HS and HS/NS binaries.
We conclude that quark-hadron stars may or may not obey these ``hadronic" universal relations, depending on the location of transition, and the difference between constituent masses in the binary.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{hybrid.eps}
\end{center}
\caption{
Fractional difference from the improved binary Love relations fit for a representative set of constrained nuclear matter EoSs (gray cirlces), and 30 selected NS binaries from the ACB5 hybrid EoS.
In particular, these binary pairs are selected from each of the following categories: (i) binaries with which one massive star has a quark matter core, while the other is pure hadronic matter (i.e. small enough mass that the transition point has not yet been reached), denoted HS/NS (green triangles); (ii) binary with both small-mass hadron stars, denoted NS/NS (red squares); and (iii) binary with both large-mass stars with quark matter cores, denoted HS/HS (magenta diamonds).
}
\label{fig:hybrid}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%% Begin Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Discussion}\label{sec:conclusion}
The recent GW observation of binary NS merger GW170817 placed constraints on the supranuclear matter EoS for NSs.
We take advantage of this by generating a restricted set of spectral EoSs which agree with this observation to reduce uncertainties upon the extraction of tidal parameters from future GW events.
Important previous work by Yagi and Yunes~\cite{Yagi:ILQ,Yagi:binLove} found EoS-insensitive universal relations between symmetric and antisymmetric combinations of NS tidal deformabilities, which aid in the extraction of said tidal parameters.
We find an improvement upon these universal relations by a factor of 25\% for stars with mass below 1.7 solar masses\zc{check this number} by restricting to the EoSs constrained by GW170817.
In addition, we find that binaries consisting of one massive hyrid quark-hadron star, and one small-mass hadron star do not obey these universal relations -- explained by the large difference in $\tilde{\Lambda}$ displayed between the constituent stars.
Further, we analyze the impact of this improvement for future binary NS merger detections, as the current detector sensitivity is not yet small enough to accurately constrain $\tilde{\Lambda}$ enough for universal relations to make a difference.
We find that future interferometers Voyager, CE, and ET-D all experience statistical uncertainties upon the extraction of $\tilde{\Lambda}$ small enough to become comparable to the systemtic errors injected by using the improved universal relations, as depicted in Fig.~\ref{fig:stackedFisher} and Table~\ref{tab:variances}.
This indicates that the use and further improvement of universal relations is justified for future GW observations of binary NS merger events.

>>>>

@@

Fig.~\ref{fig:binLove} shows the improved binary Love universal relation for 3 different mass ratios: $q=0.9$, $q=0.75$, and $q=0.5$ for both constrained and unconstrained sets of EoSs.
We observe that the maximum recorded EoS variation under the constrained set of EoSs is $23.65\%$, $9.76\%$, and $2.15\%$ for mass ratios of $q=0.9$, $q=0.75$, and $q=0.50$, respectively.
This shows considerable improvement from the universalities of $\sim 50\%$, $\sim 20\%$, and $\sim 3\%$ found in Ref.~\cite{Yagi:binLove}.
In addition, the unconstrained set of EoSs show similar universality to that recorded in Ref.~\cite{Yagi:binLove}.
\begin{figure*}
\begin{center} 
\includegraphics[width=\textwidth]{binLove.eps}
\end{center}
\caption{
Binary Love universal relations shown for the constrained EoSs (left) and unconstrained EoSs (right).
In this figure, the dashed black line corresponds to the fit given by Eq.~\ref{eq:binLovefit}, while the dotted blue, dashed magenta, and solid green lines represent the 35 EoSs for mass ratios of $q=0.9$, $q=0.75$, and $q=0.50$, respectively.
The maximal EoS variation for each value of $q$ is $23.65\%$, $9.76\%$, and $2.15\%$ respectively, compared to $\sim 50\%$, $\sim 20\%$, and $\sim 3\%$ found in Ref.~\cite{Yagi:binLove}.
These values are comparable to the maximal EoS variation for unconstrained EoSs as expected, found to be $41.66\%$, $12.64\%$, and $3.83\%$.
Additionally shown is the binary Love relations for hybrid stars ACB5 and ACB7 (dash-dash-dotted curves) as discussed in Sec.~\ref{sec:hybrid}.
Observe the large deviations from the fit as the mass ratio $q$ increases, returning near the fit before the transitional pressure at large values of $\Lambda_s$.
}
\label{fig:binLove}
\end{figure*}

@@

\subsubsection{Hybrid quark-hadron stars}\label{sec:hybrid}
\begin{figure}
\begin{center} 
\includegraphics[width=\linewidth]{hybridML.eps}
\end{center}
\caption{
$M-\tilde{\Lambda}$ relations for the ACB5 and ACB7 model hybrid stars. 
Note the strong phase transition into a quark-matter branch at $\sim 1.4 \text{ M}_{\odot}$.
As described in the text, HS/HS and NS/NS binaries have individual masses chosen such that both stars are contained within one branch -- resulting in high $q$ values.
Constrastingly, HS/NS binaries have individual masses chosen from each branch resulting in small $q$-values.
}
\label{fig:hybridML}
\end{figure} 
Transitional quark-hadron matter stars undergo first-order phase transitions at pressure $P_{\text{tr}}$, where the hadronic branch departs into a quark-matter branch at a given mass transition.
We focus on the ACB5 and ACB7 models shown in Fig.~\ref{fig:hybridML}, which observe transition masses at $\sim 1.4 \text{ M}_{\odot}$~\cite{Paschalidis2018}.
These result in two distinct types of NSs, based on their observed mass: (i) massive ($M \geq 1.4 \text{ M}_{\odot}$) hybrid stars which have quark-matter cores and nuclear matter crusts (we denote this as ``HS"), and (ii) small-mass ($M \leq 1.4 \text{ M}_{\odot}$) hadronic stars with no internal transition to quark matter (we denote this as ``NS").
We consider 3 classes of binaries:
\begin{itemize}
\item HS/NS binary: consists of one massive hybrid star, and one small hadronic neutron star,
\item NS/NS binary: consists of two small-mass hadronic neutron stars,
\item HS/HS binary: consists of two massive hybrid stars.
\end{itemize}

@@

To determine how each of class of binary NSs obey our improved binary Love universal relations, we pick a representative set of 10 binary pairs from each class.
Binary Love values are then computed for each pair, then residuals from the binary Love curve found in Sec.~\ref{sec:nuclear} are determined and compared to that from the constrained set of EoSs.
The results are shown in Fig.~\ref{fig:hybrid}.
First, we observe that the NS/NS combination of stars gives consistent results to that found earlier -- as expected.
Next, we see that HS/HS combinations agree moderately with previous relations\footnote{Note that these EoSs do not fit within GW170817's 90\% posterior on EoS - so we excpect deviations consistent with the ``unconstrained" set of EoSs, rather than the ``constraind" set.}, while HS/NS combinations disagree significantly.
This can be explained by the large reduction in tidal deformability for hybrid stars, are shown in Fig. 2 of Ref.~\cite{Paschalidis2018}: when one star lies on the hadronic branch while the other is on the quark branch, they each see vast differences in $\tilde{\Lambda}$ - disrupting the binary Love relations as found for stars on the hadronic branch.
In addition, the discrepancy can be seen in Fig.~\ref{fig:binLove} where the binary Love relations for the ACB5 and ACB7 EoSs are shown in relation to the constrained and unconstrained nuclear matter EoSs.
Observe how the residuals from the fit are quite good for low pressures (large $\Lambda_s$) when the the binaries consist of a NS/NS combination, before reaching the transitional pressure when the deviations become increasingly large for HS/HS and HS/NS binaries.
We conclude that quark-hadron stars may or may not obey these ``hadronic" universal relations, depending on the location of transition, and the difference between constituent masses in the binary.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{hybrid.eps}
\end{center}
\caption{
Fractional difference from the improved binary Love relations fit for a representative set of constrained nuclear matter EoSs (gray cirlces), and 30 selected NS binaries from the ACB5 hybrid EoS.
In particular, these binary pairs are selected from each of the following categories: (i) binaries with which one massive star has a quark matter core, while the other is pure hadronic matter (i.e. small enough mass that the transition point has not yet been reached), denoted HS/NS (green triangles); (ii) binary with both small-mass hadron stars, denoted NS/NS (red squares); and (iii) binary with both large-mass stars with quark matter cores, denoted HS/HS (magenta diamonds).
}
\label{fig:hybrid}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%% Begin Discussion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Discussion}\label{sec:conclusion}
The recent GW observation of binary NS merger GW170817 placed constraints on the supranuclear matter EoS for NSs.
We take advantage of this by generating a restricted set of spectral EoSs which agree with this observation to reduce uncertainties upon the extraction of tidal parameters from future GW events.
Important previous work by Yagi and Yunes~\cite{Yagi:ILQ,Yagi:binLove} found EoS-insensitive universal relations between symmetric and antisymmetric combinations of NS tidal deformabilities, which aid in the extraction of said tidal parameters.
We find an improvement upon these universal relations by a factor of 25\% for stars with mass below 1.7 solar masses\zc{check this number} by restricting to the EoSs constrained by GW170817.
We also find an increase in I-Love-Q universality by a factor of 23\%.
In addition, we find that binaries consisting of one massive hyrid quark-hadron star, and one small-mass hadron star do not obey these universal relations -- explained by the large difference in $\tilde{\Lambda}$ displayed between the constituent stars.
Further, we analyze the impact of this improvement for future binary NS merger detections, as the current detector sensitivity is not yet small enough to accurately constrain $\tilde{\Lambda}$ enough for universal relations to make a difference.
We find that future interferometers Voyager, CE, and ET-D all experience statistical uncertainties upon the extraction of $\tilde{\Lambda}$ small enough to become comparable to the systemtic errors injected by using the improved universal relations, as depicted in Fig.~\ref{fig:stackedFisher} and Table~\ref{tab:variances}.
This indicates that the use and further improvement of universal relations is justified for future GW observations of binary NS merger events.
